---
title: 'HW4: Nonlinear and Ensemble  Modeling (due 12/12 Monday 23:30)'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Problem 1: Bike Sharing Dataset

This data set collects the bike sharing counts aggregated on daily basis (2011-2012) together with the calender information (year/season/weekday/holiday) and daily weather conditions. 
There are 4 weather realted variables:  

* weathersit: (GOOD, MISTY, RAIN/SNOW/STORM) 

* temperature (in Celsius)  

* humidity (relative to 100%)

* windspeed 

The objective is to predict the bike sharing counts given a date, and figure out the impacts on the counts due to weather conditions.

More data descriptions can be found at http://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset

### Your analysis should include the following:

* EDA and preliminary findings.

* Fit tree-based models and summarize their prediction performance. The tuning parameters for tree-based models should be determined via cross validation.

* Fit a linear regression and a nonlinear regression model for the target variable **count**. Again, your models should go through a proper model selection procedure. Summarize the selected models with their prediction performance.

* Identify the important input variables for predicting bike sharing counts. In particular, characterize the effects of weather conditions on the bike counts **quantitatively**.

* Give a brief analysis summary.


```{r}
bike <- read.csv(file="bike.csv") #read data
head(bike)
#View(bike)  
```



------------------------------------------------------------- 
## Problem 2: Airline Customer Satisfaction


This dataset contains an airline passenger satisfaction survey. 
The data collect the demographic information about the customers, their feedback regarding the flight experiences and some flight information.

The goal of the analysis has two folds: 
  
* Find a good model for the classification of customer satisfaction, and summarize the performance. 

* Identify the important aspects of the flight experience related to customer satisfaction (or dis-satisfaction).
Based on your findings, give some specific suggestions (provided with the data evidence) to the airline company for achieving a higher customer satisfaction. 

* Add some noise factors (both continuous variable and categorical variable) in your data set (call these added variables z1-z?), and check the ranking of variable importance of these z-variables relative to the original meaningful input variables. Make your comments regarding this experiment. 


More data descriptions can be found at 
https://www.kaggle.com/teejmahal20/airline-passenger-satisfaction

```{r}
survey <- read.csv(file="airline.csv") #read data
head(survey)
survey1 <- na.omit(survey) #remove missing data
```

